Holomorphic field theories and higher algebra

### 2024-10-02 draft.tex
keep on trucking

\owen{mention some history (e.g., Manin, Donaldson-Thomas, Nekrasov) and some recent work (CPTVV, Pridham, Kapranov, etc) but then explain our focus: higher analogs of VOAs and finding mathematical analogs of physics}

\owen{Notation to explain: dg = differential graded, PDE = partial differential equations}

@@

A community has formed, involving both mathematicians and physicists, to explore and exploit these relationships,
and it continues to grow as there is a lot to explore here.
Examples of efforts in this direction include \cite{PTVV}, \cite{FHK}, \cite{Keviny things}, \cite{GaiottoEtAl}, \cite{Paquette}, \cite{Vicedo}, \cite{BeemBZEtAl}, 
where our own background and interests (and unfortunately, our ignorance) have shaped this list.

@@

@@

In physics, a classical field theory is usually described in terms of a few pieces of data:
\begin{itemize}
\item a manifold $M$ that plays the role of ``spacetime,''
\item a space of fields, such as sections of some fiber bundle $F \to M$ or connections on a principal bundle on $M$, and
\item a system of PDE for the fields, known as the equations of motion.
\end{itemize}
These equations of motion are variational: 
they are the Euler-Lagrange equations of an action functional on the space of fields,
so that solutions to these PDE are the critical points, in a sense, of some function.
In practice an action functional is a formal expression, 
written as an integral over $M$ of some polydifferential operator on the fields that takes values in densities of~$M$.
A few examples will illuminate this sketch.

@@

There are holomorphic versions of all these basic types, 
and we now spell out some simple examples.\footnote{See \cite{Donaldson Thomas} for a pioneering and illuminating discussion of how to find such analogs.}
To give the readers --- possibly a mixed audience of mathematicians and physicists --- some feel for the examples, 
we describe them in several ways: 
in terms of PDE and action functionals but also as moduli spaces.
We do not offer a thorough justification of the dictionary, for which see \cite{CosSUSY} for a wonderful and motivating exposition,
and see \cite{AlfonsoYoung} and \cite{Steffens} for recent work on this topic.

@@

A recent survey of factorization algebras, with a focus on their role in mathematical physics, can be found in~\cite{CosGwEMP}.

>>>>

@@

A community has formed, involving both mathematicians and physicists, to explore and exploit these relationships,
and it continues to grow as there is a lot to explore here.
Examples of efforts in this direction include \cite{PTVV}, \cite{FHK}, \cite{Keviny things}, \cite{GaiottoEtAl}, \cite{Paquette}, \cite{Vicedo}, \cite{BBBDN}, 
where our own background and interests (and unfortunately, our ignorance) have shaped this list.

@@

As complementary perspectives to that offered here, we recommend
\begin{itemize}
\item Kapranov's lectures \cite{KapLect} on factorization algebras in algebraic geometry,
\item Costello's lectures \cite{CosScheim} on how holomorphic field theories emerge from supersymmetric theories, and
\item a recent survey \cite{GKW} for physicists on how higher algebra appears in perturbative QFT, with a strong emphasis on the holomorphic examples.
\end{itemize}
Reading these would give a more well-rounded view of the subject and its possibilities.

@@

In physics, a classical field theory is usually described in terms of a few pieces of data:
\begin{itemize}
\item a manifold $M$ that plays the role of ``spacetime,''
\item a space of fields, such as sections of some fiber bundle $F \to M$ or connections on a principal bundle on $M$, and
\item a system of partial differential equations (PDE) for the fields, known as the equations of motion.
\end{itemize}
These equations of motion are variational: 
they are the Euler-Lagrange equations of an action functional on the space of fields,
so that solutions to these PDE are the critical points, in a sense, of some function.
In practice an action functional is a formal expression, 
written as an integral over $M$ of some polydifferential operator on the fields that takes values in densities of~$M$.
A few examples will illuminate this sketch.

@@

There are holomorphic versions of all these basic types, 
and we now spell out some simple examples.\footnote{See \cite{Donaldson Thomas} for a pioneering and illuminating discussion of how to find such analogs. Nekrasov and collaborators also \owen{Say something and cite 1990s stuff.}}
To give the readers --- possibly a mixed audience of mathematicians and physicists --- some feel for the examples, 
we describe them in several ways: 
in terms of PDE and action functionals but also as moduli spaces.
We do not offer a thorough justification of the dictionary, for which see \cite{CosSUSY} for a wonderful and motivating exposition,
and see \cite{AlfonsoYoung} and \cite{Steffens} for recent work on this topic.

@@

Kapranov gave a lecture series explaining how factorization algebras fit into algebraic geometry \cite{KapLect}.
A recent survey of factorization algebras, with a focus on their role in mathematical physics, can be found in~\cite{CosGwEMP}.
